Astrophysical calibration of gravitational-wave detectors

### 2016-02-08 docs/response_to_referee.txt

We would like to thank the referee for their very useful comments.
We have included response to them inline below and make note where this
has resulted in changes to the manuscript. These changes are displayed in red
in the resubmitted version of the manuscript.

Major comment: "In section 5, I have a problem with the assumption of the sGRB jet
opening angle being modelled as a half-normal distribution. Figure 3
in Fong et al. (2014) shows that the long GRB distribution may be
modelled by something like a Rayleigh distribution (or something
similar to give a tail to high values), but with little weight at
small angles. If the same is true of the sGRB population (and from
Figure 3 the long tail seems to be present here too) then a
half-normal distribution for the opening angle is not a good match to
the data. Did you try other priors on the inclination angle which may
better represent the true distribution? I ask that you try to use a
better prior for the inclination angle, or alternatively provide me
with a better justification of your current choice."

Response: Our poor description regarding how we use the opening angle (theta_j)
distribution to constrain the distribution on the orientation angle (iota) of
the systems has caused confusion here. We are not really interested
in the jet opening angle distribution, but are just trying to somehow use
it as a proxy for getting an idea of the orientation angle distribution,
which would not be expected to look the same. Under the assumption that to
observe a sGRB we must have theta_j/2 >= |iota|, and assuming a Rayleigh
distribution on theta_j, the distribution on |iota| does become a half-normal
distribution. In fact our choice of a half-normal for |iota| with a sigma value of
14.8 degs is rather conservative* (in that it extends to larger iota values
[this is something we would plan to correct in future studies]) compared to the
half-normal distribution that you truly get based a Rayleigh distribution in 
theta_j with a median of 10 degs. We have added text to the manuscipt to
hopefully make this clearer.

*this conservative nature is in part due to mistakenly assuming the opening
angle referred to the half width of the beam rather than the full beam width.

@@

Comment: ""Observations of sGRBs give an inferred rate of sGRB producing CBC
mergers" -> mention that the sGRB rate must be corrected for beaming
effects to map to CBC merger rate."

Response: This has now been clarified in the text.

Comment: ""may be possible to identify the host galaxy of the source" -> is a
host galaxy necessary for redshift identification? Can the redshift
not be attained from spectroscopic analysis of the sGRB light curve?"

Response: No, unfortunately the redshift cannot be otained via the sGRB itself
since it contains no specific spectral features to allow redshift
determination, e.g. transmission or absorbtion lines.

@@

Comment: ""the sky position of the source α and δ." -> I assume these
correspond to RA and DEC, but this should be made clear."

Response: alpha and delta have now been defined as right ascension and declinarion respectively.

Comment: ""for the black hole we use a prior based on the canonical mass
distribution used for the rate results in [34] with a mean of 5 M and
a standard deviation of 1 M." -> you mention that this distribution is
only an example for this type of system, but how would you do this in
reality? Would you make the mean and standard deviation further
parameters in your search in a hierarchical modelling scheme? How
would your current choices bias calibration attempts if the population
you are targeting is significantly different than your assumptions
here?"

Response: Our choices of signal parameter prior primarily effect the ensemble
statistics, i.e. the distributions of calibration uncertainties at given
distances. These give the reader an indication of the likely accuracy of an
astrophysical calibration scheme. We specifically discuss the effect of higher
mass BH systems on our results just before Section VIA.  However, in that
discussion we do not mention bias but rather the effect of different choices of
source population where the appropriate prior would be used. With regards to
bias, in "reality" we would first hope to be able to make a calibration
measurement for single detections in which case the true source distribution is
unknown to us. In this case our assumed mass priors would be dominated by the
information content of the likelihood due to the mass dependent phasing
information.  Hence for individual detections we believe that any choice of
conservative priors will have negligible effect on the result. If in the future
one were to attempt this with a population of detections then it would be
necessary to assume a hierachical scheme with some form for the mass
distribution as the referee suggests.  However, we feel that this is beyond the
scope of this work.

@@

Response: We have added the additional text stating "the weakest signals that
could still be detected would always have a single detector \ac{SNR} of $\sim
5.5$ no matter their distance, hence the plateau.

The issue of the V1 result not saturating is addressed in the manuscript in the
lines "The largest SNR contribution will generally come from the two aLIGO
detectors and thus the detection criteria (SNR threshold) will not apply to the
AdV result. Hence, the SNR in AdV can be small and thus the ability to
constrain its scale factor becomes poor (although it still provides information
that the calibration is not grossly inaccurate)."

Comment: "In figure 1 the tick labels for the chirp mass are overlaid with
the label. Better labelling or figure formatting is needed here, as
well as units where appropriate. I like the different edge coloring
for the relevant scale factor boxes, but perhaps these red edges could
be thickened for black and white printing."

Response: We have ammended the plots to produce non-overlapping make the chirp
mass tick labels, give the chirp mass units, and thicker red borders on the
scale factor subplots.

### 2016-02-08 docs/cosmic_cal.tex
Near final changes to the referee responses.

% discuss the inclination and opening angle
Other information relevant for a coincident \ac{sGRB} is that the system is
likely to be relatively close to face-on (and therefore nearly circularly
polarised), which allows us to place prior constraints on the system
inclination angle. Beam opening angles for \acp{sGRB} are relatively poorly
constrained as they are based on only a few jet-break detections. A comparative
study of estimates and lower-limits of opening angles for 13 \acp{sGRB} given
in~\cite{2014ApJ...780..118F} provides a median opening angle value of $\sim
10^{\circ}$. \refresp{We use this median opening angle as a rather conservative
guide to form a Gaussian prior distribution on the system inclination angle,
$\iota$, by converting it into an equivalent standard devaition for a half-normal
distribution. We therefore place a Gaussian prior on $\iota$ with zero mean and a standard deviation
$\sigma=14.8^{\circ}$.} This, and the choice of polarisation angle prior range given below, 
explicitly fixes the handedness of the system's rotation (or alternatively whether it is face-on or 
back-on). If we were to allow the inclination to have two modes (one for face-on and one 
for back-on), or alternatively to double the range of the polarisation angle, it would have no 
effect on our ability to estimate the signal amplitude as the two modes should be identical with 
regards to the amplitude parameter probability distributions. However, we are consistent between the 
range of simulated signals discussed in Sec.~\ref{sec:simulations} and the initial ranges stated.

>>>>

% discuss the inclination and opening angle
Other information relevant for a coincident \ac{sGRB} is that the system is
likely to be relatively close to face-on (and therefore nearly circularly
polarised), which allows us to place prior constraints on the system
inclination angle. Beam opening angles for \acp{sGRB} are relatively poorly
constrained as they are based on only a few jet-break detections. A comparative
study of estimates and lower-limits of opening angles for 13 \acp{sGRB} given
in~\cite{2014ApJ...780..118F} provides a median opening angle value of $\sim
10^{\circ}$. \refresp{We use this median opening angle as a rather conservative
guide to form a Gaussian prior distribution on the system inclination angle,
$\iota$, by converting it into an equivalent standard deviation for a
half-normal distribution. We therefore place a Gaussian prior on $\iota$ with
zero mean and a standard deviation $\sigma=14.8^{\circ}$.} This, and the choice
of polarisation angle prior range given below, explicitly fixes the handedness
of the system's rotation (or alternatively whether it is face-on or back-on).
If we were to allow the inclination to have two modes (one for face-on and one
for back-on), or alternatively to double the range of the polarisation angle,
it would have no effect on our ability to estimate the signal amplitude as the
two modes should be identical with regards to the amplitude parameter
probability distributions. However, we are consistent between the range of
simulated signals discussed in Sec.~\ref{sec:simulations} and the initial
ranges stated.
